Collaborative Test-Driven Scientific Model Validation

### 2013-11-22 sciunit-icse14/sciunit-icse14.tex
New abstract

%\usepackage{biblatex}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{times}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{setspace}
\usepackage{python_syntax}
%\usepackage{bibentry}
\usepackage{todonotes}
\usepackage{url}
\usepackage{lineno}
%\usepackage{natbib}
%\usepackage{caption}
\usepackage[english]{babel} 
\def\UrlFont{\em} % Italicize all URLs.  
\lstset{
	basicstyle=\small\ttfamily,
	breaklines=true,
	language=Python,
	}
\let\verbx\lstinline
\newcommand{\bfhead}[1]{\noindent \textbf{#1:}}

%\usepackage{alltt}
%\renewcommand{\ttdefault}{txtt}
%%\usepackage{color}
%%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%%\definecolor{mauve}{rgb}{0.58,0,0.82}
%\definecolor{light-gray}{gray}{0.75}
%\usepackage{listings}
%\lstset{
%  language=Python,
%  showstringspaces=false,
%  formfeed=\newpage,
%  tabsize=4,
%  commentstyle=\itshape,
%  basicstyle=\ttfamily\scriptsize,
%  morekeywords={lambda, self, assert, as},
%  numbers=left,
%  numberstyle=\scriptsize\color{light-gray}\textsf,
%  xleftmargin=2em,
%  stringstyle=\color{mauve}
%}
%\lstdefinestyle{Bash}{
%    language={}, 
%    moredelim=**[is][\color{blue}\bf\ttfamily]{`}{`},
%}
%\lstdefinestyle{OpenCL}{
%	language=C++,
%	morekeywords={kernel, __kernel, global, __global, size_t, get_global_id, sin}
%}
%
%\usepackage{float}
%\floatstyle{ruled}
%\newfloat{codelisting}{tp}{lop}
%\floatname{codelisting}{Listing}

@@

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations
\numberofauthors{2}
\author{\alignauthor
Cyrus Omar, Jonathan Aldrich\\\affaddr{Carnegie Mellon University, Pittsburgh, PA}\\
       \email{\{comar,aldrich\}@cs.cmu.edu}
       \alignauthor Rick Gerkin\\
       \affaddr{Arizona State University}\\{rgerkin@asu.edu}
}

@@

\begin{abstract}

Rigorously validating a scientific model's explanatory power requires comparing its predictions against empirical data on an ongoing basis. 
However, model validation remains an informal and incomplete process, especially in fields like neuroscience where the available data is growing rapidly.   
This has made it difficult to develop a contemporary understanding of a model's strengths and weaknesses, to adequately compare competing models, to determine the state-of-the-art, and to precisely identify open modeling problems.

Software engineers validate software by writing simple executable tests, called ``unit tests'', that individually check a portion of a program against a single correctness criterion. 
Suites of such unit tests can collectively provide evidence for its validity and correctness. 
Drawing inspiration from this practice, we develop the \textit{SciUnit} framework for developing suites of ``model validation tests'' -- 
executable functions, here written in Python, that compare model predictions against a single empirical observation to produce a score indicating agreement between the model and the data. 
Suites of such validation tests are used as a summary of modeling progress within a scientific community. 
As a first application of this framework, we describe \textit{NeuronUnit}, a library that supports validation testing for neurophysiology and integrates with several existing neuroinformatics resources.

%in the context of the applications and architectures that the language may need to target.
%
%Programming languages targeted toward end-users in high-performance computing must consider a number of issues in addition to performance. Typically, researchers have focused verifiability, portability and ease-of-use, and the right balance often depends on the particular uses that they are put toward.
%
%Ace %, evaluates it against several design and adoption criteria, and describes several use cases as well as a realistic case study involving parallelizing neurobiological circuit simulations on clusters of GPUs. 
%is an extensible statically-typed programming language embedded within the ubiquitous scripting language Python. Python serves as a {metalanguage}, enabling programmatic control over several aspects of language specification and compilation. 
%We demonstrate the power of this design for HPC by defining all of the primitives of the OpenCL kernel language as a library of user-defined types in Ace, then extending this core with several higher-level first-class abstractions. The target of translation, called the \emph{backend}, is user-defined as well, allowing us to implement the semantics of these primitives by direct translation to OpenCL, CUDA or C99.
%
%Although Ace is built around a static type discipline, it supports a novel form of generic programming based on type propagation that, together with a form of extensible local type inference, eliminates most of the syntactic overhead usually associated with statically-typed languages. A companion host API allows type information from \verb|numpy| arrays to directly propagate into Ace kernels, supporting seamless adoption of Ace into current scientific workflows.
%
\end{abstract}

@@

% Section 2.
\section{Validation Testing with \textit{SciUnit}}
\subsection{Example: The Quantitative Single Neuron Modeling Competition}
\begin{figure}
\small
\begin{python}
class SpikeCountTest(sciunit.Test):
  """Tests spike counts produced in response to several current stimuli against observed means and standard deviations.

>>>>

%\usepackage{biblatex}
%\usepackage{amsmath}
%\usepackage{amssymb}
%\usepackage{times}
\usepackage{hyperref}
%\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}
%\usepackage{setspace}
\usepackage{python_syntax}
%\usepackage{bibentry}
\usepackage{todonotes}
\usepackage{url}
\usepackage{lineno}
%\usepackage{natbib}
%\usepackage{caption}
\usepackage[english]{babel} 
\def\UrlFont{\em} % Italicize all URLs.  
\lstset{
%	basicstyle=\tiny\ttfamily,
	breaklines=true,
	language=Python,
	}
\let\verbx\lstinline
\newcommand{\bfhead}[1]{\noindent \textbf{#1:}}

\usepackage{alltt}
\renewcommand{\ttdefault}{txtt}
%%\usepackage{color}
%%\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
%%\definecolor{mauve}{rgb}{0.58,0,0.82}
%\definecolor{light-gray}{gray}{0.75}
%\usepackage{listings}
%\lstset{
%  language=Python,
%  showstringspaces=false,
%  formfeed=\newpage,
%  tabsize=4,
%  commentstyle=\itshape,
%  basicstyle=\ttfamily\scriptsize,
%  morekeywords={lambda, self, assert, as},
%  numbers=left,
%  numberstyle=\scriptsize\color{light-gray}\textsf,
%  xleftmargin=2em,
%  stringstyle=\color{mauve}
%}
%\lstdefinestyle{Bash}{
%    language={}, 
%    moredelim=**[is][\color{blue}\bf\ttfamily]{`}{`},
%}
%\lstdefinestyle{OpenCL}{
%	language=C++,
%	morekeywords={kernel, __kernel, global, __global, size_t, get_global_id, sin}
%}
%
%\usepackage{float}
%\floatstyle{ruled}
%\newfloat{codelisting}{tp}{lop}
%\floatname{codelisting}{Listing}

@@

% author names and affiliations
% use a multiple column layout for up to two different
% affiliations
\numberofauthors{2}
\author{\alignauthor
Cyrus Omar, Jonathan Aldrich\\\affaddr{Carnegie Mellon University}\\
       \email{\{comar,aldrich\}@cs.cmu.edu}
       \alignauthor Richard C. Gerkin\\
       \affaddr{Arizona State University}\\{rgerkin@asu.edu}
}

@@

\begin{abstract}
One of the pillars of the modern scientific method is \emph{model validation}: comparing a scientific model's predictions against empirical observations. Today, a scientist demonstrates the validity of a model by making an argument in a paper and submitting it for peer review, a process comparable to \emph{code review} in software engineering. While peer review ensures that contributions meet high-level goals, software engineers typically supplement this process with \emph{unit testing} to get a more complete picture of the state of a software project, particularly for complex projects involving many developers.

We argue that a similar test-driven methodology would be valuable to scientific communities that need to validate increasingly complex models against growing collections of empirical data. The dynamics of scientific communities differ from software communities in several key ways, however. In this paper, we introduce \emph{SciUnit}, a test-driven model validation framework designed to handle these differences in a practical manner. We also describe how SciUnit, supported by both new and existing collaborative infrastructure, can be integrated into the modern scientific process.
%in the context of the applications and architectures that the language may need to target.
%
%Programming languages targeted toward end-users in high-performance computing must consider a number of issues in addition to performance. Typically, researchers have focused verifiability, portability and ease-of-use, and the right balance often depends on the particular uses that they are put toward.
%
%Ace %, evaluates it against several design and adoption criteria, and describes several use cases as well as a realistic case study involving parallelizing neurobiological circuit simulations on clusters of GPUs. 
%is an extensible statically-typed programming language embedded within the ubiquitous scripting language Python. Python serves as a {metalanguage}, enabling programmatic control over several aspects of language specification and compilation. 
%We demonstrate the power of this design for HPC by defining all of the primitives of the OpenCL kernel language as a library of user-defined types in Ace, then extending this core with several higher-level first-class abstractions. The target of translation, called the \emph{backend}, is user-defined as well, allowing us to implement the semantics of these primitives by direct translation to OpenCL, CUDA or C99.
%
%Although Ace is built around a static type discipline, it supports a novel form of generic programming based on type propagation that, together with a form of extensible local type inference, eliminates most of the syntactic overhead usually associated with statically-typed languages. A companion host API allows type information from \verb|numpy| arrays to directly propagate into Ace kernels, supporting seamless adoption of Ace into current scientific workflows.
%
\end{abstract}

@@

% Section 2.
\section{Validation Testing with \textit{SciUnit}}
\subsection{Example: The Quantitative Single Neuron Modeling Competition}
\begin{figure}[t]
\small
\begin{python}
class SpikeCountTest(sciunit.Test):
  """Tests spike counts produced in response to several current stimuli against observed means and standard deviations.
